---
title: "MFA with mfaMKTLT"
author: "Minchul Shin, Ko Donghyeon, Taehee Jung, Lev Golod, Temi N Lal"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with mfaMKTLT}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(collapse = T, comment = "#>")
library(mfaMKTLT)
```

## Acknowledgements

The authors of the package would like to thank Professor Gaston Sanchez, as  
well as Herve Abdi1, Lynne J. Williams, and Domininique Valentin (WIREs Comput  
Stat 2013. doi: 10.1002/wics.1246)  

-------------------------------------------------------------------------------

## When to use MFA?

MFA (multifactor analysis) is a techinque designed to "handle multiple data  
tables that measure sets of variables collected on the same observations." For  
example, the file `winedata` which comes included with this package, has the  
results of a fictitious wine-tasting experiment, where 10 wine critics rated  
12 different wines from different regions. 

-------------------------------------------------------------------------------

## Preliminaries

Let's familiarize ourselves a bit with this data set. Each wine critic used  
4 common variables (V1...V4) to rate the wines, and was also free to choose up  
to two additional variables. Whenever we see "V1" or "V1." in the variable  
names, that indicates the beginning of a new block or sub-table. For example,  
the first 6 columns correspond to critic 1, and the next 6 correspond to critic  
2. The different wines are represented as row names.  

In general, when using the `"mfaMKTLT"` package, the data must be either a data  
frame or a matrix.  
```{r prelim1}
head(winedata[1:12])
is.data.frame(winedata)
```

-------------------------------------------------------------------------------

## Sets

As mentioned above, the different critics' assessments correspond to different  
sub-tables or blocks in our data. To use the `mfa()` function, we must supply  
`sets`, a list of variable names or indices that dictates how the data are  
organized into sub-tables. `sets` should be a list where each element is  
either a character vector or a numeric vector. All of the following are valid:
```{r sets1}
sets.num <- list(c(1:6), c(7:12), c(13:18), c(19:23), c(24:29), c(30:34),
  c(35:38), c(39:44), c(45:49), c(50:53))
sets.char <- list(c('V1','V2','V3','V4','V5','V6'),
  c('V1.1','V2.1','V3.1','V4.1','V7','V8'),
  c('V1.2','V2.2','V3.2','V4.2','V9','V10'),
  c('V1.3','V2.3','V3.3','V4.3','V8.1'),
  c('V1.4','V2.4','V3.4','V4.4','V11','V12'),
  c('V1.5','V2.5','V3.5','V4.5','V13'),
  c('V1.6','V2.6','V3.6','V4.6'),
  c('V1.7','V2.7','V3.7','V4.7','V14','V5.1'),
  c('V1.8','V2.8','V3.8','V4.8','V15'),
  c('V1.9','V2.9','V3.9','V4.9'))
sets.mixed <- c(sets.num[1:4], sets.char[5:10])
```

Note that a given variable (column) may only belong to one sub-table. Note also  
that not every variable in the data must be used. Only the variables that  
belong to a sub-table are considered 'active variables.' In the wine data, the  
following variables do not correspond to the assessment of any critic, and are  
therefore inactive:
```{r sets2}
all.equal(unlist(sets.num), unique(unlist(sets.num)))
all.equal(unlist(sets.char), unique(unlist(sets.char)))
head(winedata[c("acidity", "ph", "alcohol", "res_sugar")])
```

-------------------------------------------------------------------------------

## Default Function Arguments

By default `ncomps = NULL` means that the MFA conducts a Generalized  
Singular Value Decomposition and returns all of the Principal Components. The  
user can choose a different value for `ncomps` if she or he wishes (any integer  
between 1 and L, where L is the rank of the data).  

By default `center = TRUE` means that each variable (column) is centered by  
dividing by its own mean.  

By default `scale = 'vector.norm'` means that each variable is scaled by  
dividing by its vector norm (Euclidean length). Note that all scaling is done  
after centering is completed.  

-------------------------------------------------------------------------------

## Creating an `"mfa"` Object  

Let's create an `"mfa"` object using the wine data, and the 'mixed' sets list  
created earlier. We will use the default values for `ncomps`, `center`, and  
`scale`. If we print `mfa1`, we will see a brief summary about how the analysis  
was conducted. 
```{r mfa1}
mfa1 <- mfa(data = winedata, sets = sets.num)
mfa1
```

An `"mfa"` object is a list, so we can subset its various elements using  
dollar-sign subsetting. For an explanation of each element, see the help  
documentation for `mfa()`. 
```{r mfa2}
names(mfa1)
help(mfa, package = "mfaMKTLT")
```

Some of the most important elements of the output are:  

|            |                                                  |
|:-----------|:-------------------------------------------------|
| `eigvals`  | The vector of eigenvalues.                       |
| `Fcommon`  | The matrix of common factor scores.              |
| `Fpartial` | A list of matrices of partial factor scores.     |
| `Q`        | The matrix of loadings (right singular vectors). |

Let's take a quick look at each of these elements.  
```{r mfa3}
head(mfa1$eigvals)
head(mfa1$Fcommon[,1:5])
head(mfa1$Fpartial[[1]][,1:5])
head(mfa1$Q[,1:5])
```

-------------------------------------------------------------------------------

## Summary Methods

The `"mfaMKTLT"` package provides a variety of summary methods to help study  
different aspects of the analysis. `eigsum` provides a __summary of eigenvalues__,  
singular values, and inertia.  
```{r sum1}
eigsum <- eigsum(mfa1)
round(eigsum[, 1:5], 3)
```

-------------------------------------------------------------------------------

We can look at the __contributions__ of observations, variables, and tables to  
components, with the functions `obscont`, `varcont`, and `tabcont`  
respectively.  
```{r sum2}
# The [i,j] element is the contribution of observation i to component j.
obscont <- obscont(mfa1) 
round(obscont[1:5, 1:5], 3) 

# The [i,j] element is the contribution of variable i to component j.
varcont <- varcont(mfa1) 
round(varcont[1:5, 1:5], 3) 

# The [i,j] element is the contribution of sub-table i to component j.
tabcont <- tabcont(mfa1) 
round(tabcont[1:5, 1:5], 3) 
```

-------------------------------------------------------------------------------

There are two ways to consider __between-table relationships__: The Rv  
coefficient and the Lg coefficient both tell us how similar two sub-tables are.  
The Rv can take values in `[0-1]` while Lg can take any positive real value.  
For both, higher values indicate a higher degree of similarity. 

Let's consider the Rv and Lg coefficients between sub-tables 2 and 5.  
```{r sum3}
rv_coeff(mfa1$subtable[[2]], mfa1$subtable[[5]])
lg_coeff(mfa1$subtable[[2]], mfa1$subtable[[5]])
```

If we are interested in studying the between-table structure of multiple  
sub-tables at once, we can use the `gentable` function to see Rv or Lg  
coefficients. The [i,j] element of the resulting matrix tells us about the  
similarity of sub-tables i and j. Note that the matrix is symmetric.  

Let's consider the between-table structure for sub-tables 1-3. We will first  
display the Rv coefficients, and then the Lg.
```{r sum4}
# rv coefficients
gentable(mfa1$data_scale, as.list(c(sets.num[1], sets.char[2:3])),
          paste0("sub-table_", 1:3), func = rv_coeff)

# lg coefficients
gentable(mfa1$data_scale, as.list(c(sets.num[1], sets.char[2:3])),
          paste0("sub-table_", 1:3), func = lg_coeff)
```


